---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# predicttornado

<!-- badges: start -->
<!-- badges: end -->

The goal of predicttornado is to provide functions that help with the data analysis of predicting the number of fatalities using length and width of tornadoes based on tornado data. 

## Installation

You can install the development version of predicttornado like so:

``` r
# install.packages("devtools")
devtools::install_github("DSCI-310-2024/predicttornado")
```

## Example

Refer to the [vignettes](https://github.com/DSCI-310-2024/predicttornado/blob/main/vignettes/predicttornado-vignette.Rmd) to view more examples on how to use our package functions. 

```{r}
raw_df <- data.frame(
  om = c(156, 189, 26), 
  yr = c(1950, 1951, 1952), 
  mo = c(10, 11, 12), 
  dy = c(1, 9, 20), 
  date = as.Date(c("2021-03-26", "2022-03-26", "2024-03-01")),
  time = as.POSIXct(c("12:00:00", "13:00:00", "21:00:00"), format = "%H:%M:%S"),
  tz = c("America/Chicago", "America/Chicago", "America/Chicago"),
  datetime_utc = as.POSIXct(c("2021-03-26 12:00:00", "2022-03-26 13:00:00", "2024-03-01 21:00:00")),
  st = c("OK", "NC", "TX"),
  stf = c(0, 0, 0), 
  mag = c(0, 0, NA), 
  inj = c(23, 233, 405), 
  fat = c(45, 190, 201), 
  loss = c(23, 40, 56), 
  slat = c(0, 0, 0), 
  slon = c(0, 0, 0), 
  elat = c(0, 0, 0), 
  elon = c(0, -50, -60), 
  len = c(201, 340, 100), 
  wid = c(34, 57, 71), 
  ns = c(0, 0, NA), 
  sn = c(0, 0, 0), 
  f1 = c(0, 0, 0), 
  f2 = c(0, 0, 0), 
  f3 = c(0, 0, NA), 
  f4 = c(0, 0, 0), 
  fc = c(FALSE, FALSE, TRUE)
)
```

This is a basic example which shows you how to load the package and use one of its functions, `process_data`, which processes and cleans your data before using it for data analysis. 

```{r example}
library(magrittr)
library(predicttornado)

tornado_df <- process_data(raw_df)
```

This package can also be used to create visualizations like scatterplots using `create_scatterplot`and boxplots using `boxplot_viz`. 

```{r}
create_scatterplot(tornado_df, width, fatalities) + 
ggplot2::labs(x = "Width of tornadoes (yards)", y = "Fatalities", 
title = "Figure 2: Scatterplot of width (yards) of tornado and fatalities")
```

```{r}
boxplot_viz(tornado_df, fatalities)
```

You can easily create a linear regression model and fit it to your data with the function `fit_linear_model`: 

```{r}
lm_data <- fit_linear_model(fatalities ~ width + length, tornado_df) |>
    predict(tornado_df) |>
    dplyr::bind_cols(tornado_df)
```

## predicttornado in the R package ecosystem 

`predicttornado` may seem like a package very similar to other R packages. For example, we have three visualization functions that resemble `ggplot2`. However, they are all slightly different and more tailored to our data analysis project.  `create_scatterplot`and `boxplot_viz` have specific visualization size settings and title and aesthetics that render your visualizations more pleasing. 

Moreover, `accuracy_plot` is very different because even though it uses `ggplot2` functions, it creates a scatterplot that compares the accuracy of a linear regression model to the actual observations. 

`process_data` is also different from `dplyr`'s `select` and `filter` as it combines these different functions to create an efficient data processing function for ease of data analysis. Additionally, it is a specific processing function for analysis tornado data in predicting fatalities. 

Lastly, `fit_linear_model` is also different from other linear regression functions such as `lm` from `stats` for example because it combines multiple steps required in creating and fitting a linear regression model and then automatically outputs useful regression metrics. 

 
